---
title: "R Language Basics"
output: html_notebook
---

## Simple Calculations in R, Calling Functions, and Getting Help in R
```{r}
4+3
```
```{r}
4-3
```
```{r}
4*3
```
```{r}
4/3
```

R has numerous mathematical functions. For example, we call the function sqrt() on a numeric argument:
```{r}
sqrt(3.5)
```

## Getting help in R
Each of R's functions has integrated documentation accessible within R. 
We can access R's built-in documentation with the help() function or its syntactic shortcut, ?:
```{r}
help(log)
```
```{r}
?log
```

## Variables and Assigment
To store a value for future use, we assign it to a variable using the <- assignment operator:
```{r}
x <- 3.1
```

Variables can also be used in expressions and passed to functions as arguments. R will substitute the variable's value during evaluation and return the results.
```{r}
(x+2)/2
```

When we assign a value in our R session, we're assigning it to an environment known as the global environment.
We can see objects we've created in the global environment with the function ls():
```{r}
ls()
```

## Vectors, Vectorization, and Indexing
A vector is a container of contiguous data.
Unlike most languages, R does not have a type for a single value such as 3.1 or "AGCTACGACT". Rather, these valuse are stored in a vector of length 1.
We can verify that values like 3.1 are vectors of length 1 by calling the function length():
```{r}
length(3.1)
```

To create longer vectors, we combine values with the function c():
```{r}
x <- c(56, 95.3, 0.4)
y <- c(3.2, 1.1, 0.2)
```

R's vector are the basis of one of R's most important features: vectorization.
Vectorization allows us to loop over vectors elementwise, without the need to write an explicit loop.
```{r}
x + y
```

Integer sequences crop up all over computing and statistics, so R has a few ways of creating these vectors.
```{r}
seq(3, 5)
```
```{r}
1:5
```

There's one important subtle behavior of vectorized operations applied to two vectors simultaneously: if one vector is longer than the other, R will recycle the values in the shorter vector.
R won't watn you whrn this happens (unless the recycled shorter vector's length isn't a multiple of the longer vector's length).

```{r}
x - 3
```

```{r}
x/2
```

```{r}
c(1,2) + c(0, 0, 0, 0)
```

```{r}
c(1,2) + c(0, 0, 0)
```

Many of R's mathematical functions are all vectorized:
```{r}
sqrt(x)
```
```{r}
round(sqrt(x), 3)
```
```{r}
log(x)/2 + 1
```

