---
title: "Indexing"
output: html_notebook
---

We can access specific elements of a vector through indexing.
An index is an integer that specifies which element in the vector to retrieve.
```{r}
x <- c(56, 95.3, 0.4)
x[2]
```

R's vectors are 1-indexed, meaning that the index 1 correspnds to the first element in a list.
```{r}
x[1]
```

Trying to access an element that doesn't exist in the vector leads R to return NA, the "not available" missing value.
```{r}
x[4]
```

We can change specific vector elements by combining indexing and assignments.
```{r}
x[3] <- 0.5
```

Vectors can also have names, you can set with c():
```{r}
b <- c(a = 3.4, b = 5.4, c = 0.4)
b
```

Access names of a vector:
```{r}
names(b)
```

Change names:
```{r}
names(b) <- c("x", "y", "z")
```

Access elements by their name:
```{r}
b['x']
```
```{r}
b['y']
```
```{r}
b['z']
```

Subsetting:
```{r}
x[c(2, 3)]
```

Extracting contiguous slices of a vector:
```{r}
z <- c(3.4, 2.2, 0.4, -0.4, 1.2)
z[3:5]
```

Exclude certain elements from lists using negative indexes:
```{r}
z[c(-4, -5)]
```

Indices are also often used to reorder elements.
Rearrage the elements of this vector z with:
```{r}
z[c(3, 2, 4, 5, 1)]
```

Reverse the elements of this vector:
```{r}
z[5:1]
```

We can use other R functions to create indexes for us.
For example, the function order() returns a vector of indexes that indicate the (ascending) order of the elements.
```{r}
order(z)
```
```{r}
z[order(z)]
```
```{r}
order(z, decreasing = TRUE)
```
```{r}
z[order(z, decreasing = TRUE)]
```

Repeat certain vaues in vectors:
```{r}
z[c(2, 2, 1, 4, 5, 4, 3)]
```

Resample a vector is to randomly sample its indexes uding the sample() function:
```{r}
set.seed(0)
i <- sample(length(z), replace = TRUE)
i
```

```{r}
z[i]
```

Use comparison operatoes like ==, !=, <, <=, >, and >= to builf logical vectors of TRUE and FALSE values indicating the result of the comparison test for each element in the vector.
```{r}
v <- c(2.3, 6, -3, 3.8, 2, -1.1)
v == 6
```


Indexing vectors with logival vectors
```{r}
v[c(TRUE, TRUE, FALSE, TRUE, FALSE, FALSE)]
```

Subset by creating logival vectors using comparison operators:
```{r}
v[v>2]
```

```{r}
v > 2 & v < 4
```

```{r}
v[v > 2 & v < 4]
```













