---
title: "Vector types"
output: html_notebook
---

R's vectors must contain elements of the same type.
R supports the following vector types:

- Numeric
- Integer
- Character
- Logical
- Complex (represent complex numbers)
- raw (encode raw bytes)

## Type coercion in R
R would coerce the logical TRUE and FALSE values to 1 and 0, as these represent TRUE and FALSE without a loss of informtaion:
```{r}
c(4.3, TRUE, 2.1, FALSE)
```

```{r}
c(-9L, 3.4, 1.2, 3L)
```

```{r}
c(43L, TRUE, 3.2413341, "a string")
```

We can see any object's type using the function typeof():
```{r}
q <- c(2, 3.5, -1.1, 3.8)
typeof(q)
```

## Factors 
Factors stored categorical variables, such as a treatment group, strand, or chromosome.

Create a factor from vector using the function factor():
```{r}
chr_hits <- c("chr2", "chr2", "chr3", "chrX", "chr2", "chr3", "chr3")
hits <- factor(chr_hits)
hits
```

The levels are the possible values a factor can contain. 
View a factor's levels by using the function levels():
```{r}
levels(hits)
```

When creating our factor, we could use the argument levels to include all relevant levels:
```{r}
hits <- factor(chr_hits, levels = c("chrX", "chrY", "chr2", "chr3", "chr4"))
hits
```

Add or rename existing levels with the function levels().
```{r}
levels(hits) <- list(chrX = "chrX", chrY = "chrY", chr2 = "chr2", chr3 = "chr3", chr4 = "chr4")
hits
```

Count up how many of each level there are in a factor:
```{r}
table(hits)
```

