---
title: "Working with and Visualizing Data in R"
output: 
  md_document:
    variant: markdown_github
---

## Loading Data into R
```{r}
d <- read.csv("Dataset_S1.txt")
```

## Exploring and Transforming Dataframes
```{r}
head(d, n=3)
```

```{r}
nrow(d)
```

```{r}
ncol(d)
```

```{r}
dim(d)
```

```{r}
colnames(d)
```

change the 12th column name
```{r}
colnames(d)[12]
```
```{r}
colnames(d)[12] <- "percent.GC"
```
```{r}
colnames(d)[12]
```

The most common way to access a single column of a dataframe is with the dollar sign operator.
```{r}
d$depth
```

```{r}
mean(d$depth)
```

```{r}
summary(d$depth)
```

To access the first two columns 
```{r}
d[,1:2]
```

Similarly, we could use the column names
```{r}
d[,c("start", "end")]
```

If we only wanted the first row of start and end positions, we'd use:
```{r}
d[1, c("start", "end")]
```

Let's add an additional column to our dataframe that indicates whether a window is in the centromere region.
We can append to d dataframe using comparison and logical operations:
```{r}
d$cent <- d$start >= 25800000 & d$end <= 29700000
```

